Mergeable Types

### 2017-02-28 paper/implementation.tex
Edits.

\begin{wrapfigure}{L}{.5\textwidth}
	\begin{center}
	\includegraphics[scale=0.8]{Figures/partitions}
	\end{center}
	\caption{Network partitions may leave branches $b_1$ and $b_4$ in one
		partition (red background), and branches $b_2$ and $b_3$ in the other
		(yellow background). Since $b_2$ and $b_3$ cannot merge unless $b_1$ and
		$b_4$ merge, they cannot communicate among themselves even though they are in
		the same network partition.}
	\label{fig:partitions}
\end{wrapfigure}

Concurrent operations in Irmin are tracked within the notion of branches,
allowing the programmer to explicitly merge the branches on demand. \name
realizes the concurrent programming model on top of the notion of branches;
each replica operates on its own branch, and thus is isolated from actions on
other replicas. In presence of network partitions, the nodes in one partition
may not receive updates from the other partition, hence merges cannot happen
across a partition. This can be a problem if a pair of branches in one
partition (the current partition) have multiple LCAs in the other partition
(the remote partition), hence cannot merge. Fig.~\ref{fig:partitions}
illustrate this problem for the branching structure of
Fig.~\ref{fig:external-lcas}.

Fortunately, the access to full (causal) branching history assured by Irmin at
every node comes to the rescue. Relying on the history, the current partition
can fork-off (\C{fork}) new branches that start where the remote branches left,
and map them to new (virtual) nodes that emulate the nodes in the remote
partition. For the example in Fig.~\ref{fig:partitions}, the partition of $b_1$
and $b_4$ can fork new branches $b_2'$ and $b_3'$ from $b_2$ and $b_3$,
respectively, and resume the activity as if the partition never happened. After
merging the first versions on $b_2'$ and $b_3'$, branches $b_1$ and $b_2$ will
be in the same state as before (when the partition happened), except that the
branches are now mergeable because $b_2'$ and $b_3'$ can merge. The ability to
track full provenance information is thus crucial for \name to overcome network
partitions, making it an appropriate programming model for highly-available
replicated data types. \name monad conceals the branching structure, but also
transparently performs the necessary merges to obtain a history graph with a
unique lowest common ancestor.

>>>>

\begin{wrapfigure}{L}{.5\textwidth}
	\begin{center}
	\includegraphics[scale=0.8]{Figures/partitions}
	\end{center}
  \caption{Network partitions may leave branches $b_1$ and $b_4$ in
  one partition (red background), and branches $b_2$ and $b_3$ in the
  other (yellow background). Access to full history on both sides of
  the partition lets both sides make progress.}
	\label{fig:partitions}
\end{wrapfigure}

Concurrent operations in Irmin are tracked within the notion of branches,
allowing the programmer to explicitly merge the branches on demand. \name
realizes the concurrent programming model on top of the notion of branches;
each replica operates on its own branch, and thus is isolated from actions on
other replicas. In presence of network partitions, the nodes in one partition
may not receive updates from the other partition, hence merges cannot
happen across a partition. This can be a problem if a pair of branches
in one partition (the current partition) have their loci
(Sec.~\ref{sec:meta}) in the other partition (the remote partition),
hence cannot merge. Fig.~\ref{fig:partitions} illustrate this problem
for the branching structure of
Fig.~\ref{fig:external-lcas}.

Fortunately, the access to full (causal) branching history assured by
Irmin at every node comes to the rescue. Relying on the history, the
current partition can fork-off (\C{fork}) new branches that start
where the remote branches left, and map them to new (virtual) nodes
that emulate the nodes in the remote partition. For the example in
Fig.~\ref{fig:partitions}, the partition of $b_1$ and $b_4$ can fork
new branches $b_2'$ and $b_3'$ from $b_2$ and $b_3$, respectively, and
resume the activity as if the partition never happened. The ability to
track full provenance information is thus crucial for \name to
overcome network partitions, making it an appropriate programming
model for highly-available replicated data types. \name monad conceals
the branching structure, but also transparently performs the necessary
merges to obtain a history graph with a unique lowest common ancestor.
